Developing a Bubble Chamber Particle Discriminator Using Semi-Supervised Learning

### 2018-12-01 documentation/paper.tex
Updated information for GD

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{waveform_hist}
    \caption{\label{waveform_hist} Prediction distribution of the best {\it DeepConv}$(\omega)$ configuration. The black vertical line represents the AP decision boundary.}
\end{figure}

@@

Because all data in the PICO-60 experiment is labeled, unlabeled training data sets are randomly sampled from the 496 training examples remaining (after the validation set is selected).

@@

During a grid search of parameters of the gravitational differentiation algorithm, in addition to the stochastic gradient descent learning rate (a multiplier that defines how quickly and precisely optimization occurs), the highest mean accuracy over three runs of an individual configuration was 99\%. The highest accuracy was obtained with 256 labeled examples in the set $\varsigma$ (out of 496 total training examples), which equates to 52\% of the data being labeled. High accuracy statistics in this range were also obtained using 128 examples (99\% accuracy, once again) and 64 examples (97\% accuracy). 32 examples produced a much lower maximum accuracy, at only 83\%. From this, it is evident that it is possible to achieve very high accuracy using this technique. It is also clear that, in general, larger initial sets produce higher performance.

@@

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Gravitational differentiation accuracy statistics by size of $\varsigma$.} \label{grav_accuracy_by_varsigma}
        \begin{tabular}{|l|l|l|l|}
            \hline
            Initial $\varsigma$ Size & Proportion of Data Labeled & Max Accuracy & Mean Accuracy \\
            \hline
            32 & 6\% & 83\% & 69\% \\
            \hline
            64 & 13\% & 97\% & 73\% \\
            \hline
            128 & 26\% & 99\% & 84\% \\
            \hline
            256 & 52\% & 99\% & 94\% \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

Table \ref{grav_overview} represents the performance of several high-accuracy gravitational differentiation configurations tested during the grid search. The accuracy statistic is the mean over three tests of the same configuration.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of gravitational differentiation configurations.} \label{grav_overview}
        \begin{tabular}{|l|l|l|l|l|l|l|l|}
            \hline
            $\varsigma$ Size & $\delta g$ & Learning Rate & $\psi$ & Mean AP Accuracy & Precision & Recall & CWSD \\
            \hline
            32 & 0.0005 & 0.001 & 5 & 73\% & 72\% & 90\% & 0.78 \\
            \hline
            32 & 0.0005 & 0.003 & 9 & 83\% & 80\% & 99\% & 0.63 \\
            \hline
            64 & 0.0005 & 0.003 & 9 & 77\% & 74\% & 97\% & 0.71 \\
            \hline
            64 & 0.003 & 0.03 & 7 & 97\% & 96\% & 100\% & 0.18 \\
            \hline
            128 & 0.008 & 0.001 & 9 & 74\% & 71\% & 95\% & 0.70 \\
            \hline
            128 & 0.0005 & 0.03 & 11 & 99\% & 99\% & 99\% & 0.17 \\
            \hline
            256 & 0.005 & 0.01 & 11 & 97\% & 96\% & 99\% & 0.25 \\
            \hline
            256 & 0.001 & 0.03 & 3 & 99\% & 99\% & 100\% & 0.10 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{grav_grid_search}
    \caption{\label{grav_grid_search} Predictions of best gravitational differentiation model compared to AP (on randomly selected validation set).}
\end{figure}

>>>>

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{waveform_hist}
    \caption{\label{waveform_hist} Prediction distribution of the best {\it DeepConv}$(\omega)$ configuration, relative to AP, on a randomly selected validation set. The black vertical line represents the AP decision boundary.}
\end{figure}

@@

Because all data in the PICO-60 experiment is labeled, unlabeled training data sets are randomly sampled from the 496 training examples remaining (after the validation set is selected). Labeled set sizes of 32, 64, 128, and 256 were tested; the remainder of the 496 examples would be used as unlabeled data.

@@

During a grid search of parameters of the gravitational differentiation algorithm, in addition to . The highest accuracy was obtained with 256 labeled examples in the set $\varsigma$ (out of 496 total training examples), which equates to 52\% of the data being labeled. High accuracy statistics in this range were also obtained using 128 examples (99\% accuracy, once again) and 64 examples (97\% accuracy). 32 examples produced a much lower maximum accuracy, at only 83\%. From this, it is evident that it is possible to achieve very high accuracy using this technique. It is also clear that, in general, larger initial sets produce higher performance.

@@

As with supervised learning techniques, a grid search was run for various parameters of the gravitational differentiation algorithm, in addition to the stochastic gradient descent learning rate (a multiplier that defines how quickly and precisely optimization occurs). Table \ref{grav_accuracy_by_varsigma} shows the mean asymptotic accuracy over the best-performing region of the hyperparameter search space, for each size of the initial labeled set $\varsigma$. Note that optimal hyperparameters vary by set size.

It is evident that larger labeled sets produce better performance. However, note that a $\varsigma$ size of 128 produces accuracy of 97.0\%, the same as the banded supervised system with all 496 labeled examples. That is, GD achieves the same accuracy with 74\% of the labels removed.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Gravitational differentiation accuracy statistics by size of $\varsigma$.} \label{grav_accuracy_by_varsigma}
        \begin{tabular}{|l|l|l|l|l|l|}
            \hline
            Initial $\varsigma$ Size & Proportion of Data Labeled & Accuracy & Precision & Recall & CWSD \\
            \hline
            32 & 6\% & 71.3\% & 70.1\% & 98.0\% & 0.73 \\
            \hline
            64 & 13\% & 79.7\% & 75.5\% & 98.6\% & 0.63 \\
            \hline
            128 & 26\% & 97.0\% & 96.3\% & 99.6\% & 0.23 \\
            \hline
            256 & 52\% & 99.2\% & 99.0\% & 99.7\% & 0.11 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}
% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples256 | grep rate0.03 | grep power3 | ./mean_ap_accuracy_across_configurations.py
% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples128 | grep rate0.03 | grep increment0.0005 | ./mean_ap_accuracy_across_configurations.py
% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples64 | grep power11 | grep rate0.001 | ./mean_ap_accuracy_across_configurations.py
% cat ../experimental_data/training_logs/2018_10_26_gravitational_ap_similarity_last_300_epochs.out | grep examples32 | grep power11 | grep increment0.001 | ./mean_ap_accuracy_across_configurations.py

Table \ref{grav_overview} represents the performance of several of the individual configurations incorporated into the mean accuracy values in Table \ref{grav_accuracy_by_varsigma}. All accuracy statistics are the mean over three tests of the same configuration.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of gravitational differentiation configurations in the best-performing region of the hyperparameter search space.} \label{grav_overview}
        \begin{tabular}{|l|l|l|l|l|l|l|l|}
            \hline
            $\varsigma$ Size & $\delta g$ & Learning Rate & $\psi$ & Accuracy & Precision & Recall & CWSD \\
            \hline
            32 & 0.001 & 0.001 & 11 & 79.0\% & 78.2\% & 94.8\% & 0.74 \\
            \hline
            32 & 0.001 & 0.003 & 11 & 59.9\% & 59.9\% & 100.0\% & 0.78 \\
            \hline
            64 & 0.005 & 0.001 & 11 & 82.8\% & 79.0\% & 99.1\% & 0.58 \\
            \hline
            64 & 0.008 & 0.001 & 11 & 73.9\% & 71.0\% & 99.6\% & 0.71 \\
            \hline
            128 & 0.0005 & 0.03 & 3 & 98.3\% & 97.7\% & 99.6\% & 0.21 \\
            \hline
            128 & 0.0005 & 0.03 & 9 & 96.6\% & 95.1\% & 99.6\% & 0.29 \\
            \hline
            256 & 0.008 & 0.03 & 3 & 99.1\% & 99.0\% & 99.5\% & 0.16 \\
            \hline
            256 & 0.001 & 0.03 & 3 & 99.7\% & 99.6\% & 99.6\% & 0.05 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{grav_grid_search}
    \caption{\label{grav_grid_search} Prediction distribution of the best gravitational differentiation configuration.}
\end{figure}
